---
title: "SarcomaPreOpBiopsy"
author: "Heejoo"
date: "`r Sys.Date()`"
output:
  html_document: 
    dev: jpeg
    df_print: paged
    fig_height: 6
    fig_width: 10
    highlight: textmate
    theme: cosmo
    toc: yes
    toc_float:
      collapsed: no
    code_folding: show
  word_document:
    toc: yes
    toc_depth: '3'
  bookdown::word_document2:
    fig_caption: yes
editor_options:
  chunk_output_type: console
always_allow_html: yes
---

<style>
   p {line-height: 2.5em;}
   pre, code {white-space:pre !important; overflow-x:scroll !important}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, message = F, warning = F)
```

## Load & Input
```{r}
library(readxl)
library(dplyr)
library(tableone)
library(DT)

setwd("C:/Users/USER/Desktop/2020-1학기/의학통계/sarcoma/biopsy")
a<-excel_sheets("sarcoma data sheet SMC 20200728.xlsx") %>% 
  lapply(function(x){read_excel("sarcoma data sheet SMC 20200728.xlsx",sheet=x,skip=2)})
b<-a[[1]] %>% 
  left_join(a[[2]],by="환자번호") %>% 
  left_join(a[[3]],by="환자번호") %>% 
  left_join(a[[4]],by="환자번호") %>% 
  left_join(a[[5]],by="환자번호") %>% 
  left_join(a[[6]],by="환자번호") %>% 
  left_join(a[[7]],by="환자번호")

b[["ECOG\r\n\r\n0/1/2/3/4"]][which(is.na(b[["ECOG\r\n\r\n0/1/2/3/4"]]))]<-"0"
b[["EBL\r\n(ml)"]]<-ifelse(b[["EBL\r\n(ml)"]]=="UK",NA,as.numeric(b[["EBL\r\n(ml)"]]))
b$Age<-as.numeric(b[["수술날짜\r\n\r\ndd-mm-yyyy"]]-b[["생년월일\r\n\r\ndd-mm-yyyy"]])/365.25

```

## Methods
```{r}
c <- b %>%
  filter(`Primary 수술여부\r\n\r\n0. Primary tumor\r\n1. Residual after incomplete resection\r\n2. Local recurrence.x` == "0" & `환자번호` != "21733889") %>% 
  filter(!(`수술 전 Biopsy\r\n\r\n0. None\r\n1. Primary site\r\n2. Local recurrence site\r\n3. Metastatic site` == 1 & `Type of needle\r\n\r\n0. Core\r\n1. FNA\r\n2. N/A\r\n3. Unknown` %in% 2:3)) %>% 
  mutate(biopsy_preop_primary = as.integer(`수술 전 Biopsy\r\n\r\n0. None\r\n1. Primary site\r\n2. Local recurrence site\r\n3. Metastatic site` == "1"),
         type_needle = `Type of needle\r\n\r\n0. Core\r\n1. FNA\r\n2. N/A\r\n3. Unknown`) %>% 
  mutate(type_needle = ifelse(type_needle == 0, "Core needle", ifelse(type_needle == 1, "FNA", "Excisional biopsy")))

out<-c %>% select(환자번호,Age,`성별\r\n\r\nM/F`,biopsy_preop_primary,type_needle)
names(out)[3]<-"Sex"; names(out)[1]<-"ID"
out$Sex <- as.factor(out$Sex)
out[out$ID == 31050857, "Sex"] <- "F"
```

##Data Collecting
```{r}
#Biopsy
a.rps<-excel_sheets("preop Bx. RPS 여부 20200712.xlsx") %>% 
  lapply(function(x){read_excel("preop Bx. RPS 여부 20200712.xlsx",sheet=x, skip=2, na = c("UK"))})
out.rps<-a.rps[[1]][c(4:6)]
names(out.rps)[1]<-"ID"; names(out.rps)[2]<-"RPS_preop"; names(out.rps)[3]<-"RPS_postop"
out<-merge(out,out.rps,by="ID")

out$liposarcoma_preop <- as.integer(c[["preOP Bx. 결과\r\n\r\n0. WD \r\n1. DD \r\n2. Pleomorphic \r\n3. LMS\r\n4. MPNST\r\n5. Solitary fibrous tumor\r\n6. PEComa\r\n7. Other"]] %in% c("0", "1", "1or 2", "2"))
out$liposarcoma_postop <- as.integer((c[["병리결과\r\n\r\n0. WD Liposarcoma\r\n1. DD Liposarcoma\r\n2. Pleomorphic Liposarcoma\r\n3. Leiomyosarcoma\r\n4. MPNST\r\n5. Solitary fibrous tumor\r\n6. PEComa\r\n7. Other"]] %in% c(0, 1, 2)) |
                                       (c[["병리결과\r\n\r\n0. WD Liposarcoma\r\n1. DD Liposarcoma\r\n2. Pleomorphic Liposarcoma\r\n3. Leiomyosarcoma\r\n4. MPNST\r\n5. Solitary fibrous tumor\r\n6. PEComa\r\n7. Other"]] == 7) &
                                       grepl("liposarcoma|Liposarcoma", c[["Other \r\n\r\ncomment"]]))
out$DDLPS_postop <- as.integer(c[["병리결과\r\n\r\n0. WD Liposarcoma\r\n1. DD Liposarcoma\r\n2. Pleomorphic Liposarcoma\r\n3. Leiomyosarcoma\r\n4. MPNST\r\n5. Solitary fibrous tumor\r\n6. PEComa\r\n7. Other"]] == 1)

#Patients survival rate-death,day-FU
out$death<-as.integer(c[["사망여부\r\n\r\n0.Alive\r\n1.Dead\r\n2.Unknown.y"]])
out$death <- ifelse(out$death == 2, NA, out$death) ## Death 2 -> NA 로 수정.
out$day_FU<-as.numeric(c[["마지막 f/u\r\n\r\ndd-mm-yyyy"]]-c[["수술날짜\r\n\r\ndd-mm-yyyy"]])

#Local recurrence free survival rate-recur_local,recur_site,recur_day
out$recur_local<-c[["재발#1\r\n\r\n0: 무\r\n1: 유.x"]]
out$recur_site<-c[["Site of local recurrence"]]
out$recur_site<-ifelse(out$recur_site=="6",NA,out$recur_site)
out$recur_day<-ifelse(out$recur_local==1,
                      as.numeric(as.Date(as.integer(c[["Date of local recurrence"]]),origin="1899-12-30")-as.Date(c[["수술날짜\r\n\r\ndd-mm-yyyy"]])),
                      as.numeric(c[["마지막 f/u\r\n\r\ndd-mm-yyyy"]]-c[["수술날짜\r\n\r\ndd-mm-yyyy"]]))
out$recur_day[is.na(out$recur_day)] <- as.numeric(as.Date("02/10/18", "%m/%d/%y") - as.Date("10/02/12", "%m/%d/%y"))

#RT-RTdose,RTx_tissue_expander
out$RTx_dose<-c[["RT dose\r\n(Gy)"]] 
cond1<-c[["RT timing\r\n\r\n0.None \r\n1.Preop only\r\n2. IORT only\r\n3.Preop + IORT\r\n4.Postop only\r\n5.Preop + postop boost\r\n6.IORT + postop"]] %in% c("1","5")
cond2<-(c[["RT timing\r\n\r\n0.None \r\n1.Preop only\r\n2. IORT only\r\n3.Preop + IORT\r\n4.Postop only\r\n5.Preop + postop boost\r\n6.IORT + postop"]] %in% "4") & (c[["Tisuue expander insertion \r\n유뮤\r\n\r\n0. No\r\n1. Yes"]]=="1")
out$RTx_tissue_expander<-as.integer(cond1|cond2)

#Neoadjuvant therapy
out$RTx_preop<-as.integer(c[["수술전 \r\nRT 여부\r\n\r\n0.No\r\n1.Yes"]])
out$Chemo_preop<-as.integer(c[["수술전 \r\nChemo 여부\r\n\r\n0.No\r\n1.Yes"]])
out$Neoadjuvant<-as.integer(out$RTx_preop|out$Chemo_preop)

out$meta_liver<-c[["Liver metastasis\r\n\r\n0. No\r\n1. Yes"]]
out$meta_liver <- ifelse(out$meta_liver == 3, NA, out$meta_liver)  ## 3인 것 발견. NA로 수정
out$meta_lung<-c[["Lung metastasis\r\n\r\n0. No\r\n1. Yes"]]
out$meta_bm<-c[["Bone metastasis\r\n\r\n0. No\r\n1. Yes"]]
out$meta_abd<-c[["Intra-abdominal metastasis\r\n\r\n0. No\r\n1. Yes"]]
out$multifocal<-c[["Mutifocality 여부\r\n\r\n0. No\r\n1. Yes"]]
out$meta<-c[["Lung metastasis\r\n\r\n0. No\r\n1. Yes"]]+
          c[["Liver metastasis\r\n\r\n0. No\r\n1. Yes"]]+
          c[["Bone metastasis\r\n\r\n0. No\r\n1. Yes"]]+
          c[["Intra-abdominal metastasis\r\n\r\n0. No\r\n1. Yes"]]
out$meta<-ifelse(out$meta==0,0,1)

#동반절제 장기수
info.resection<-c %>% 
  select(starts_with("동반절제")) %>% 
  mutate_at(1:25,as.integer) %>% 
  mutate_at(26,function(x){as.integer(!is.na(x))})

out$num_resected_organ<-rowSums(info.resection,na.rm=T)

#RT
out$RTx_total<-as.integer(c[["수술 전후 RT 여부\r\n\r\n0.No\r\n1.Yes"]])

#Chemo
out$Chemo_postop<-as.integer(c[["Adjuvant chemo 여부\r\n\r\n0.No\r\n1.Yes"]])
out$Chemo_both<-as.integer(out$Chemo_preop|out$Chemo_postop)

#Risk factor analysis for tumor recurrence
#tumor size
out$tumor_size<-c[["종양 크기\r\n(Tumor size, mm)\r\n다발성인 경우 largest tumor size"]]

#resection margin
out$resection_margin<-c[["Surgical margins\r\n\r\n0. R0/R1\r\n1. R2\r\n2. Not available"]]
out$resection_margin<-ifelse(out$resection_margin=="2",NA,out$resection_margin)

#FNCLCC tumor grade
out$FNCLCC_grade<-c[["FNCLCC grade\r\n\r\n1. total score 2-3\r\n2. total score 4-5\r\n3. total score 6,7,8"]]
out$FNCLCC_grade<-ifelse(out$FNCLCC_grade=="UK",NA,out$FNCLCC_grade)

#sarcomatosis pattern
out$sarcomatosis_pattern <- as.integer(c[["Site of local recurrence"]] == "4")
```

Variable categorizing
```{r}
my_vars<-names(out)
factor_vars <- c(names(out)[sapply(out, function(x){length(table(x))}) <= 6])
conti_vars <- setdiff(my_vars, factor_vars)
for (v in factor_vars){out[[v]] <- factor(out[[v]])}


#factor_vars 중 fisher test 해야하는 변수들
vars.fisher <- sapply(factor_vars,
                      function(x){is(tryCatch(chisq.test(table(out[["biopsy_preop_primary"]], out[[x]])),
                                              error = function(e) e, warning=function(w) w), "warning")})
vars.fisher <- factor_vars[vars.fisher]

out
datatable(cbind(c(1:27),factor_vars,c(conti_vars,rep("",20))),
          colnames = c("","factor_vars","conti_vars"))
```

### Table1 : preOP biopsy 의 정확성
```{r}
out %>% filter(biopsy_preop_primary==1) -> out.biopsy
out.biopsy$RPS_preop<-as.factor(ifelse(out.biopsy$RPS_preop=="0","No","Yes"))
out.biopsy$RPS_postop<-as.factor(ifelse(out.biopsy$RPS_postop=="0","No","Yes"))
out.biopsy$RTx_tissue_expander<-as.factor(ifelse(out.biopsy$RTx_tissue_expander=="0","No","Yes"))

table.acc<-table(out.biopsy$RPS_preop,out.biopsy$RPS_postop)
table.acc<-cbind(c("RPS_preop",""),rownames(table.acc),table.acc)
colnames(table.acc)<-c(" ","RPS_postop","Yes")
rownames(table.acc)<-NULL
datatable(table.acc)

tb <- table(out.biopsy$RPS_preop,out.biopsy$RTx_tissue_expander)
tb.prop <- round(100 * prop.table(tb, 2), 2)      ## prop.table : 1- byrow 2 - bycol
tb.out <- matrix(paste0(tb, " (", tb.prop, ")"), ncol =2)
p <- fisher.test(tb)$p.value

table1<-cbind(rownames(tb),tb.out,c(ifelse(p < 0.001, "< 0.001", round(p, 3)),""))
table1<-rbind(c("",as.integer(summary(out$RTx_tissue_expander))[1:2],""),table1)
colnames(table1)<-c("RTx_tissue_expander","No","Yes","p-value")
rownames(table1)<-c("n","RPS_preop (%)","")

datatable(table1)
```

### Table2 : preOP biopsy (+) or (-) group의 차이 확인
```{r}
table2_vars<-c("Age","Sex","tumor_size","liposarcoma_preop","liposarcoma_postop",
               "RPS_preop","RPS_postop","DDLPS_postop","type_needle",
               "FNCLCC_grade","meta_lung","meta_bm","meta_abd",
               "multifocal","num_resected_organ","RTx_total",
               "RTx_dose","RTx_tissue_expander","Chemo_both",
               "resection_margin")
table2_factor_vars<-table2_vars[table2_vars %in% factor_vars]
table2_conti_vars<-table2_vars[table2_vars %in% conti_vars]

table2<- lapply(table2_vars, function(va){
  if (va %in% table2_conti_vars){
    
    forms <- as.formula(paste0(va, "~ biopsy_preop_primary"))
    mean_sd <- aggregate(forms, data = out, FUN = function(x){c(mean = mean(x), sd = sd(x))})
    
    p<-oneway.test(forms, data = out)$p.value
    
    out.final <- c(va, "",
                   paste0(round(mean_sd[[va]][, "mean"], 2), " (", round(mean_sd[[va]][, "sd"], 2), ")"),
                   ifelse(p < 0.001, "< 0.001", round(p, 3)))
    
    return(out.final)
    
  } else if(va %in% table2_factor_vars) {
    
    tb <- table(out[[va]], out[["biopsy_preop_primary"]])
    tb.prop <- round(100 * prop.table(tb, 2), 2)      ## prop.table : 1- byrow 2 - bycol
    tb.out <- matrix(paste0(tb, " (", tb.prop, ")"), ncol =2)
    
    p <- ifelse(va %in% vars.fisher, fisher.test(tb)$p.value, chisq.test(tb)$p.value)
    
    out.final <- cbind(c(paste0(va, " (%)"), rep("", nrow(tb.out) - 1)),
                       rownames(tb),
                       tb.out,
                       c(ifelse(p < 0.001, "< 0.001", round(p, 3)), rep("", nrow(tb.out) - 1)))
    return(out.final)
  }
}) %>% Reduce(rbind, .)

table2<-rbind(c("n","",as.integer(summary(out$biopsy_preop_primary))[1:2],""),table2)
colnames(table2) <- c("Variable", "Subgroup", "No", "Yes", "p-value")
rownames(table2)<-NULL
datatable(table2)

```

### Supplement table 1. Histologic subtypes of retroperitoneal sarcoma
```{r}
```

### Table3 : pattern of recurrence after preOP biopsy (안전성)
```{r}
#other retroperitoneal site.?
#needle tract recurrence 0이지만 table에 포함시키라고 되어있음

table3_vars<-c("recur_local","meta","sarcomatosis_pattern")

table3_factor_vars<-table3_vars[table3_vars %in% factor_vars]
table3_conti_vars<-rbind(table3_vars[table3_vars %in% conti_vars])

table3<- lapply(table3_vars, function(va){
  if (va %in% table3_conti_vars){
    
    forms <- as.formula(paste0(va, "~ biopsy_preop_primary"))
    mean_sd <- aggregate(forms, data = out, FUN = function(x){c(mean = mean(x), sd = sd(x))})
    
    p<-oneway.test(forms, data = out)$p.value
    
    out.final <- c(va, "",
                   paste0(round(mean_sd[[va]][, "mean"], 2), " (", round(mean_sd[[va]][, "sd"], 2), ")"),
                   ifelse(p < 0.001, "< 0.001", round(p, 3)))
    
    return(out.final)
    
  } else if(va %in% table3_factor_vars) {
    
    tb <- table(out[[va]], out[["biopsy_preop_primary"]])
    tb.prop <- round(100 * prop.table(tb, 2), 2)      ## prop.table : 1- byrow 2 - bycol
    tb.out <- matrix(paste0(tb, " (", tb.prop, ")"), ncol =2)
    
    p <- ifelse(va %in% vars.fisher, fisher.test(tb)$p.value, chisq.test(tb)$p.value)
    
    out.final <- cbind(c(paste0(va, " (%)"), rep("", nrow(tb.out) - 1)),
                       rownames(tb),
                       tb.out,
                       c(ifelse(p < 0.001, "< 0.001", round(p, 3)), rep("", nrow(tb.out) - 1)))
    return(out.final)
  }
}) %>% Reduce(rbind, .)

table3<-rbind(table3,
              c("needle tract recurrence","0",
                paste0(summary(out$biopsy_preop_primary)[1],"(100)"),
                paste0(summary(out$biopsy_preop_primary)[2],"(100)"),"<0.001"))
table3<-rbind(c("n","",as.integer(summary(out$biopsy_preop_primary))[1:2],""),table3)
colnames(table3) <- c("Variable", "Subgroup", "No", "Yes", "p-value")
rownames(table3)<-NULL
datatable(table3)

```

### Figure1 : preOP RPS 의 효용성 patient survival, local recurrence free survival (RPS, LPS, DDLPS)
```{r}
```

### Table4 : local recurrence risk factor analysis
```{r}
```